---
title: "Rarefaction curves, estimates, and asymptotes"
author: Dani Wiley
date:   last-modified
editor: visual
description: |
  Creates dataframe and calculates rarefaction curve, aspymtote (number of estimated parasite lineages), and estimated sampling effort needed to fully sample parasite communities for gray and plumbeous vireo species. Contact dlfwiley[at]gmail.com with any questions or requests for more data.
format:
  html:
    theme:            litera
    highlight-style:  atom-one
    page-layout:      full      # article, full   # https://quarto.org/docs/output-formats/page-layout.html
    toc: true        # Enable the table of contents
    toc-title: "Contents"  # Optional: Set a custom title for the TOC
    toc-depth: 1     # Optional: Specify the heading levels to include
    toc-location: body # Optional: Show the TOC on the left (other options are "right" or "float")
    number-sections:  false
    self-contained:         false     # !!! this can cause a render error
    code-overflow:          scroll    # scroll, wrap
    code-block-bg:          true
    code-block-border-left: "#30B0E0"
    code-copy:              false     # true, false, hover a copy buttom in top-right of code block
editor_options:  
  chunk_output_type: console
---

# Load libraries

```{r}
library(iNEXT)
library(dplyr)
library(ggplot2)
library(scales)
```

# Load data

Code last updated 7/12/2025

```{r}
#Load abundance data
bird_data <- list(
  "Bell's Vireo" = c(0,1,0,1,0,10,0,0,0,0,0,0,0,0,0,0,0,0,0),  # 12 samples total - not used in this analyses
  "Gray Vireo" = c(0,0,0,0,0,0,0,0,0,0,0,0,0,0,1,0,114,1,7), # 123 samples total
  "Plumbeous Vireo" = c(1,25,1,0,14,4,2,8,1,1,1,1,1,1,1,0,0,0,0) # 62 samples total
)
set.seed(123)
#Run iNEXT for rarefaction and extrapolation
iNEXT_output <- iNEXT(bird_data, q = 0, datatype = "abundance", endpoint = 400)
iNEXT_output
#        Assemblage         Diversity  Observed Estimator        s.e.        LCL        UCL
# 1    Bell's Vireo  Species richness  3.000000  3.916667  0.78410251  3.0000000   5.453479
# 2    Bell's Vireo Shannon diversity  1.761359  2.053182  0.42530526  1.2195986   2.886765
# 3    Bell's Vireo Simpson diversity  1.411765  1.466667  0.35284027  0.7751124   2.158221
# 4      Gray Vireo  Species richness  4.000000  4.991870  0.43769275  4.1340079   5.849732
# 5      Gray Vireo Shannon diversity  1.365874  1.392621  0.09308512  1.2101775   1.575064
# 6      Gray Vireo Simpson diversity  1.159577  1.161096  0.05091690  1.0613004   1.260891
# 7 Plumbeous Vireo  Species richness 14.000000 53.846774 24.21468145 14.0000000 101.306678
# 8 Plumbeous Vireo Shannon diversity  6.380021  8.808168  2.19502783  4.5059925  13.110344
# 9 Plumbeous Vireo Simpson diversity  4.205689  4.438967  0.83133379  2.8095829   6.068351
```

##--------------------------------------------------------------

# Extracting estimated N of parasite lineages

## Bell's vireo: Nest = 3.92, s.e. 0.784, LCL = 3, UCL = 5.45

## Gray vireo: Nest = 4.99, s.e.= 0.529, LCL = 4, UCL = 6.03

## Plumbeous vireo: Nest = 53.8, s.e.= 21.0, LCL = 14, UCL = 95.1

```{r}
#EXTRACTING THE ESTIMATED TOTAL NUMBER OF PARASITE LINEAGES ------------------------------
# Extract asymptote estimates
asymptote_data <- iNEXT_output$AsyEst
parasite_rich <- filter(asymptote_data, asymptote_data$Diversity == "Species richness")
print(parasite_rich)
#        Assemblage        Diversity Observed Estimator       s.e.       LCL        UCL
# 1    Bell's Vireo Species richness        3  3.916667  0.7841025  3.000000   5.453479
# 2      Gray Vireo Species richness        4  4.991870  0.4376927  4.134008   5.849732
# 3 Plumbeous Vireo Species richness       14 53.846774 24.2146815 14.000000 101.306678

# Write out csv
#write.csv(parasite_rich, "parasite_richness_iNEXT.csv")
```

## --------------------------------------------------------------

# Extracting estimated screening effort to sample 95-99% of parasite lineages possible

## Bells vireo: Nest = 33, qD = 3.89, qDLCL = 2.01, qDUCL = 5.76

## Gray vireo: Nest = 14, qD = 1.81, qDLCL = 1.13, qDUCL = 2.49

## Plumbeous vireo: Nest = 364, qD = 40.6, qDLCL = 12, qDUCL = 69.1

```{r}
#EXTRACTING THE ESTIMATED NUMBER OF SCREENING EFFORT NEEDED -----------------------------
#Trying to add completeness (estimated n of birds needed to screen to capture)
# Extract completeness-based estimates
completeness_data <- iNEXT_output$iNextEst$coverage_based

# View the extrapolated completeness data
print(completeness_data)

# Filter for 95-99% completeness (SC >= 0.95)
sampling_effort_95_99 <- completeness_data[completeness_data$SC >= 0.95 & completeness_data$SC < 1, ]

#Add row IDs to manually select cut-offs for minimum samples needed to sample >95% parasite diversity
sampling_effort_95_99$ID <- seq.int(nrow(sampling_effort_95_99))
sampling_effort_95_99
#          Assemblage        SC   m        Method Order.q        qD    qD.LCL    qD.UCL ID
# 8      Bell's Vireo 0.9957759  33 Extrapolation       0  3.889210  1.828794  5.949626  1 *** selected ***
# 9      Bell's Vireo 0.9998505  53 Extrapolation       0  3.915695  1.814131  6.017258  2
# 10     Bell's Vireo 0.9999955  74 Extrapolation       0  3.916638  1.820713  6.012562  3
# 11     Bell's Vireo 0.9999998  94 Extrapolation       0  3.916666  1.806228  6.027103  4
# 12     Bell's Vireo 1.0000000 114 Extrapolation       0  3.916667  1.794200  6.039134  5
# 13     Bell's Vireo 1.0000000 135 Extrapolation       0  3.916667  1.789519  6.043814  6
# 14     Bell's Vireo 1.0000000 155 Extrapolation       0  3.916667  1.781384  6.051950  7
# 15     Bell's Vireo 1.0000000 175 Extrapolation       0  3.916667  1.788886  6.044447  8
# 16     Bell's Vireo 1.0000000 196 Extrapolation       0  3.916667  1.758539  6.074795  9
# 17     Bell's Vireo 1.0000000 216 Extrapolation       0  3.916667  2.057145  5.776189 10
# 310      Gray Vireo 0.9567981  14   Rarefaction       0  1.808121  1.149348  2.466893 11 *** selected ***
# 41       Gray Vireo 0.9658965  21   Rarefaction       0  2.081460  1.478640  2.684280 12
# 5        Gray Vireo 0.9715033  27   Rarefaction       0  2.271209  1.706765  2.835653 13
# 61       Gray Vireo 0.9761098  34   Rarefaction       0  2.455834  1.907916  3.003752 14
# 71       Gray Vireo 0.9791705  41   Rarefaction       0  2.613140  2.070036  3.156243 15
# 81       Gray Vireo 0.9811316  48   Rarefaction       0  2.752542  2.210663  3.294422 16
# 91       Gray Vireo 0.9821988  54   Rarefaction       0  2.862859  2.321147  3.404570 17
# 101      Gray Vireo 0.9829581  61   Rarefaction       0  2.984946  2.443232  3.526659 18
# 111      Gray Vireo 0.9833762  68   Rarefaction       0  3.102834  2.561112  3.644557 19
# 121      Gray Vireo 0.9835671  74   Rarefaction       0  3.202042  2.660276  3.743808 20
# 131      Gray Vireo 0.9836765  81   Rarefaction       0  3.316693  2.774927  3.858460 21
# 141      Gray Vireo 0.9837209  88   Rarefaction       0  3.430800  2.889028  3.972572 22
# 151      Gray Vireo 0.9837357  95   Rarefaction       0  3.544699  3.002924  4.086474 23
# 161      Gray Vireo 0.9837391 101   Rarefaction       0  3.642274  3.100499  4.184049 24
# 171      Gray Vireo 0.9837398 108   Rarefaction       0  3.756098  3.214322  4.297873 25
# 181      Gray Vireo 0.9837398 115   Rarefaction       0  3.869918  3.328143  4.411694 26
# 19       Gray Vireo 0.9837398 117   Rarefaction       0  3.899224  3.357449  4.441000 27
# 20       Gray Vireo 0.9840021 123      Observed       0  4.000000  3.458205  4.541795 28
# 21       Gray Vireo 0.9842601 124 Extrapolation       0  4.015998  3.474181  4.557815 29
# 22       Gray Vireo 0.9874647 138 Extrapolation       0  4.214680  3.673548  4.755812 30
# 23       Gray Vireo 0.9901778 153 Extrapolation       0  4.382895  3.850049  4.915742 31
# 24       Gray Vireo 0.9921776 167 Extrapolation       0  4.506880  3.961861  5.051898 32
# 25       Gray Vireo 0.9938707 182 Extrapolation       0  4.611851  4.073582  5.150119 33
# 26       Gray Vireo 0.9951186 196 Extrapolation       0  4.689221  4.152848  5.225594 34
# 27       Gray Vireo 0.9961751 211 Extrapolation       0  4.754726  4.218772  5.290680 35
# 28       Gray Vireo 0.9969538 225 Extrapolation       0  4.803007  4.267285  5.338730 36
# 29       Gray Vireo 0.9976131 240 Extrapolation       0  4.843885  4.308522  5.379248 37
# 30       Gray Vireo 0.9980991 254 Extrapolation       0  4.874014  4.339105  5.408922 38
# 31       Gray Vireo 0.9985105 269 Extrapolation       0  4.899523  4.365171  5.433874 39
# 32       Gray Vireo 0.9988138 283 Extrapolation       0  4.918324  4.384498  5.452151 40
# 33       Gray Vireo 0.9990705 298 Extrapolation       0  4.934242  4.400937  5.467548 41
# 34       Gray Vireo 0.9992598 312 Extrapolation       0  4.945975  4.413103  5.478847 42
# 35       Gray Vireo 0.9994200 327 Extrapolation       0  4.955909  4.423430  5.488387 43
# 36       Gray Vireo 0.9995381 341 Extrapolation       0  4.963230  4.431057  5.495403 44
# 37       Gray Vireo 0.9996380 356 Extrapolation       0  4.969429  4.437521  5.501336 45
# 38       Gray Vireo 0.9997117 370 Extrapolation       0  4.973998  4.442290  5.505706 46
# 39       Gray Vireo 0.9997741 385 Extrapolation       0  4.977866  4.446330  5.509403 47
# 40       Gray Vireo 0.9998230 400 Extrapolation       0  4.980897  4.449494  5.512301 48
# 381 Plumbeous Vireo 0.9517678 364 Extrapolation       0 40.558813 10.911888 70.205739 49 *** selected ***
# 391 Plumbeous Vireo 0.9548238 382 Extrapolation       0 41.400718 11.271701 71.529734 50
# 401 Plumbeous Vireo 0.9576860 400 Extrapolation       0 42.189281 11.603586 72.774975 51

host_sampling_effort_95_99 <- subset(sampling_effort_95_99, 
                                     sampling_effort_95_99$ID == "1" | 
                                     sampling_effort_95_99$ID == "11" |
                                     sampling_effort_95_99$ID == "49")
print(host_sampling_effort_95_99)
#          Assemblage        SC   m        Method Order.q        qD    qD.LCL    qD.UCL ID
# 8      Bell's Vireo 0.9957759  33 Extrapolation       0  3.889210  1.828794  5.949626  1
# 310      Gray Vireo 0.9567981  14   Rarefaction       0  1.808121  1.149348  2.466893 11
# 381 Plumbeous Vireo 0.9517678 364 Extrapolation       0 40.558813 10.911888 70.205739 49

write.csv(host_sampling_effort_95_99, "host_sampling_effort_95_99.csv")
```

#Visualization

```{r}
#PLOT RAREFACTION CURVES--------------------------------------------
# Define custom colors for each host species; #slightly darker than original colors
custom_colors <- c(
  "Bell's Vireo" = "#E3C50F",  
  "Gray Vireo" = "#8A8A8A",   
  "Plumbeous Vireo" = "#65A2E5" 
)
#This plot shows
# 
# Solid lines: Observed rarefaction.
# Dashed lines: Extrapolated diversity.
# Shaded areas: 95% confidence intervals.
# Each host species represented as a separate curve

p <- ggiNEXT(iNEXT_output, type = 1, se=TRUE) +
  scale_color_manual(values = custom_colors) +
  scale_fill_manual(values = custom_colors) + # Match fill to line colors
  scale_shape_manual(values=c(19, 19, 19)) +
  theme_minimal() +
  labs(
    title = "Rarefaction and extrapolation curves",
    x = "Number of Infections Sampled",
    y = "Number of Parasite Lineages"
  ) +
  scale_x_continuous(limits=c(0,380), #set number of axis ticks
                     expand = c(0,0), #removes within-plot margin
                     breaks = scales::pretty_breaks(n = 10))  + 
  scale_y_continuous(limits=c(0,80), #set axis limits
                     expand = c(0,0), #removes within-plot margin
                     breaks = scales::pretty_breaks(n = 10)) + #set number of axis ticks
  theme_bw() + 
  theme(plot.title.position = "plot", plot.title = element_text(face="bold")) +
  theme(legend.position = "none") +
  theme(axis.text.y=element_text(size=10), 
        axis.text.x=element_text(size=10), 
        axis.title=element_text(size=12)) +
  theme(plot.background = element_blank()) + 
  theme(panel.grid.major = element_blank()) + 
  theme(panel.grid.minor = element_blank()) + 
  theme(panel.background = element_blank()) + 
  theme(panel.background = element_blank()) +
  theme(legend.position = "blank")
plot(p)
ggsave(p, filename="Fig2.pdf", bg="transparent", height=6, width=8, units="in")
```
